---
title: "p8105_hw2_yl4604"
author: "Yu"
date: "September 25, 2020"
output: github_document
---

##Problem 1

```{r}
library(tidyverse)
library(readxl)
```


Import and clean the trash wheel dataset

```{r}
trashwheel_df = 
  read_xlsx('./Trash-Wheel-Collection-Totals-8-6-19.xlsx',
            sheet = 'Mr. Trash Wheel', 
            range = cell_cols('A:N')) %>% 
            janitor::clean_names() %>% 
            drop_na(dumpster) %>% 
            mutate(sports_balls = round(sports_balls)) %>%
            mutate(sports_balls = as.integer(sports_balls))

```

Import and clean data for 2017 and 2018
```{r}
df_2017 =
  read_xlsx('./Trash-Wheel-Collection-Totals-8-6-19.xlsx', 
            sheet = '2017 Precipitation', 
            range = 'A2:B14') %>% 
            janitor::clean_names() %>% 
            mutate(year = 2017)

df_2018 =
  read_xlsx('./Trash-Wheel-Collection-Totals-8-6-19.xlsx', 
            sheet = '2018 Precipitation', 
            range = 'A2:B14') %>% 
            janitor::clean_names() %>% 
            mutate(year = 2018)
  
```
 Join the two datasets
 
```{r}
df_2017_monthname = 
  mutate(df_2017, month_name = month.name)
df_2018_monthname = 
  mutate(df_2018, month_name = month.name)
df_2017_and_2018 = 
  bind_rows(df_2017_monthname, df_2018_monthname)
```
 
This dataset contains information from the Mr.TRASHWHEEL TRASH COLLECTOR in Baltimore, Maryland. An trash enters the inner harbor, the trashwheel collects that trash, and sotre it in a sumpster. The dataset contaisn information on year, month, and trash collected, includes some specific kinds of trash.
There are a total of `r nrow(trashwheel_df)` rows in our final dataset. 
Additional data sheets include month precipitation data.


##Problem 2

Import and clean the dataset

```{r}
nyc_transit_df = 
  read_csv('./NYC_Transit_Subway_Entrance_And_Exit_Data.csv') %>% 
  janitor::clean_names() %>% 
  select(c(line:entry, vending, ada)) %>% 
  mutate(entry = recode(entry, `YES` = TRUE, `NO` = FALSE)) %>% 
  relocate(ada, vending)

```

This dataset contains information related to each entrance of NYC Transit. In the cleaning steps, I converted the variables names to more reasonable ones, and selected out variables that this problem focuses on. At the last step, I transfered the type of entry variable from character to logical.
The resulting dataset's has `r nrow(nyc_transit_df)` rows and `r ncol(nyc_transit_df)`columns.
The dataset is not tidy now because routes served is spread across 11 columns, which correspond to 11 different routes.

There are total `r nrow(distinct(nyc_transit_df, line, station_name))` distinct stations.   

Filter out stations that are not ADA compliant.
```{r}
nyc_df_ada = 
  filter(nyc_transit_df, ada == TRUE)
```

There are total `r nrow(distinct(nyc_df_ada, line, station_name))` stations are ADA compliant

Filter stations entrances/exits without vending
```{r}
nyc_df_novending = 
  filter(nyc_transit_df, vending != 'YES')
# 1.stations? distinct stations or just stations?
```

Filter entrances.exits that allow entrance but do not have vending
```{r}
nyc_enter_novending = 
  filter(nyc_df_novending, entry == TRUE)
```

There are `r round(nrow(nyc_enter_novending)/nrow(nyc_df_novending)*100, 2)`% station entrances/exits without vending allow entrance.


Reformatting data
```{r}
nyc_longer_df = nyc_transit_df %>% 
  mutate_at(vars(route8:route11), funs(as.character)) %>%  
  #mutate(route8 = as.character(route8), route9 = as.character(route9), route10 = as.character(route10), route11 = as.character(route11)) %>% 
  pivot_longer(route1:route11, 
               names_to = 'route_number',
               values_to = 'route_name') 
  
# filter route A 
nyc_a_df = 
  filter(nyc_longer_df, route_name == 'A')

# filter ADA compliant

nyc_a_ada = 
  filter(nyc_a_df, ada == TRUE)

```

There are total `r nrow(distinct(nyc_a_df, line, station_name))` distinct stations serve the A train.
Of the stations that serves the A train, `r nrow(distinct(nyc_a_ada, line, station_name))` are ADA compliant.


##Problem 3

Import and clean data in pols-month

```{r}

polsmonth_df = 
  read_csv('./fivethirtyeight_datasets/pols-month.csv') %>% 
  separate(mon, into = c('year','month','day'), sep = '-' ) %>% 
  mutate(month = as.numeric(month)) %>%
  mutate(month = month.abb[month]) %>% 
  mutate(president = case_when(
                    prez_gop == 0 ~ 'dem', 
                    prez_gop == 1 ~ 'gop',
                    prez_gop == 2 ~ 'gop')) %>% 
  select(-c(prez_gop, prez_dem, day))
```

Import and clean data in snp

```{r}
month_df = tibble(
  month = 1:12,
  month_name = month.abb
)
snp_df = 
  read_csv('./fivethirtyeight_datasets/snp.csv') %>% 
  separate(date, into = c('month','day','year'), sep = '/' ) %>% 
  mutate(month = as.numeric(month)) %>% 
  left_join(month_df, by = 'month') %>% 
  select(-month, -day) %>%  
  rename(month = month_name) %>% 
  relocate(month)

```

Import and clean unemployment data

```{r}
unemployment = 
  read_csv('./fivethirtyeight_datasets/unemployment.csv') %>%
  pivot_longer(Jan:Dec, 
               names_to = 'month',
               values_to = 'unemployment_rate') %>% 
  relocate(month) %>% 
  janitor::clean_names() %>% 
  mutate(year = as.character(year))

  
```

Join the datasets!

```{r}
joined_df = polsmonth_df %>% 
  left_join(snp_df, by = c('year','month')) %>% 
  left_join(unemployment, by = c('year','month')) %>% 
  mutate(year = as.numeric(year))


```

The first dataset polsmonth_df contains numbers of national politicians who are democratic or republican during the time period of Jan/1947-Jun/2015.
The second dataset snp_df contains closing values of Standard & Poors stock market index through the time perios of Jan/1950 - Jul/2015.
The third dataset unemolyment contains unemployment rate through the time period of Jan/1948 - Dec/2015


The dimension of the resulting dataset is (`r dim(joined_df)`), it contains observations through (`r range(pull(joined_df, year))`) . The names of the variables in the resulting dataset are: `r names(joined_df)`